---
title: "MLM workshop for PICUP"
author: "Chelsea Helion"
date: "6/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lme4)
library(lmerTest)
library(misty)
library(Rmisc)
library(effects)
library(arm)
library(sjPlot)
library(lattice)
library(RColorBrewer)
library(ggthemes)
```

## Load the data as a tibble (a type of data frame)

```{r echo = FALSE}
picup <- read_csv("/Users/tua37526/Desktop/picupClean2.csv")
head(picup)
```

## let's make sure the social reduction variables are correlated with one another (to make sure that combining them is ok) 
```{r}
## look at the correlation between reduceSocialSelf1 and reduceSocialSelf2
cor(picup$reduceSocialSelf1, picup$reduceSocialSelf2, use = "pairwise.complete.obs") 
## .669, not bad!
picup$reduceSocialSelfMean <- (picup$reduceSocialSelf1 + picup$reduceSocialSelf2)/2
```

## cool, that's great! Now let's run a standard linear regression with covidWorry as the predictor variable and reduceSocialSelfMean as the dependent variable
```{r}
lm1 <- lm(reduceSocialSelfMean ~ covidWorry, data = picup)
summary(lm1)
plot(effect("covidWorry", lm1), grid = TRUE)
coef(lm1)
```

```{r}
## run the model
mlm1 <- lmer(reduceSocialSelfMean ~ covidWorry + (1|coded_id), data = picup)
display(mlm1)
summary(mlm1)
confint(mlm1, oldNames = FALSE)

## calculate the ICC based on model output
mlm1_ICC <- (.72)/(.72+.78)

## examine the coefficients
coef(mlm1)
fixef(mlm1)
ranef(mlm1)
se.fixef(mlm1)
se.ranef(mlm1)

## plot the data
plot_model(mlm1, type = "re")
plot_model(mlm1, type = "pred")
plot_model(mlm1, type = "slope")
plot(mlm1)

# fit a subset of the data (just makes it easier to see)
picup2 <- picup[1:33,]
mlm2 <- lmer(reduceSocialSelfMean ~ covidWorry + (1|coded_id), data = picup2)
display(mlm2)
summary(mlm2)

# create slopes and intercepts variables based on the mlm we ran on the subset
intercepts <- coef(mlm2)$coded_id[,1]
slopes <- coef(mlm2)$coded_id[,2]

#plot the model!
 ggplot(picup2, aes(x=covidWorry, y=reduceSocialSelfMean, color=as.factor(coded_id))) + 
   geom_point(shape=20) + 
   geom_abline(intercept = intercepts, slope = slopes)

 ## test the model against the null model
 
 # create the null model
 null.mlm <- lmer(reduceSocialSelfMean ~ (1|coded_id), data = picup)
 display(null.mlm)
 
 # test our model against the null model
 anova(null.mlm, mlm1)
```

## ok, now let's run a varying slope model
```{r}

## run the model
mlm2 <- lmer(covidWorry ~ wave + (1 + wave|coded_id), data = picup)
display(mlm2)
summary(mlm2)
confint(mlm2, oldNames = FALSE)

## examine the coefficients
coef(mlm2)
se.fixef(mlm2)
se.ranef(mlm2)

## plot the data
plot_model(mlm2, type = "re")
plot_model(mlm2, type = "pred")
plot_model(mlm2, type = "slope")
plot(mlm2)

picup2 <-
  picup %>% 
  drop_na(covidWorry)

mlm3 <- lmer(covidWorry ~ wave + (1 + wave|coded_id), data = picup2)
display(mlm3)

# create slopes and intercepts variables based on the mlm we ran on the subset
intercepts <- coef(mlm3)$coded_id[,1]
slopes <- coef(mlm3)$coded_id[,2]

## plot it!
  nb.cols = 101
  MLMCols <- colorRampPalette(brewer.pal(9, "Blues"))(nb.cols)
  picup2$Predictions <- predict(mlm2)
  ggplot(aes(x = wave, y = Predictions, color = as.factor(coded_id)), data = picup2) +
    geom_line(size=.5) +
    geom_point(aes(y = covidWorry)) +
    geom_abline(slope= -.37, intercept=3.45, lwd= 1.5, col = 'red') +
    scale_x_continuous("Study Wave",c(0:2)) +
    xlab("Study Wave") +
    ylab("COVID-19 related worry") +
    ggthemes::theme_calc() +
    scale_color_manual(values = MLMCols) +
    theme(legend.position = "none")
  
## test the model against the null model
 
 # create the null model
 null.mlm <- lmer(covidWorry ~ (wave|coded_id), data = picup)
 display(null.mlm)
 
 # test our model against the null model
 anova(null.mlm, mlm2)
```

## If we have time, we'll look at interactions
```{r}
## run the model
mlm3 <- lmer(covidWorry ~ wave * age_check + (wave|coded_id), data = picup)
display(mlm3)
summary(mlm3)
confint(mlm3, oldNames = FALSE)

## I go on a small tirade about mean-centering
picup$age_check_cb <- center(picup$age_check)

## run the model
mlm4 <- lmer(covidWorry ~ wave *age_check_cb + (wave|coded_id), data = picup)
display(mlm4)
summary(mlm4)
confint(mlm4, oldNames = FALSE)

## I continue talking about how great mean-centering is
picup$age_check_19 <- (picup$age_check - 19)
picup$age_check_53 <- (picup$age_check - 53)

## run the model
mlm5 <- lmer(covidWorry ~ wave *age_check_19 + (wave|coded_id), data = picup)
display(mlm5)
mlm6 <- lmer(covidWorry ~ wave *age_check_53 + (wave|coded_id), data = picup)
display(mlm6)

## plot the data
plot_model(mlm4, type = "re")
plot_model(mlm4, type = "pred")
plot(effect("wave:age_check", mlm3), grid = TRUE)

## examine the coefficients
coef(mlm3)
se.fixef(mlm3)
se.ranef(mlm3)

## examine the coefficients
coef(mlm4)
se.fixef(mlm4)
se.ranef(mlm4)


## test the model against the null model
 
 # create the null model
 null.mlm <- lmer(covidWorry ~ wave + age_check_cb + (wave|coded_id), data = picup)
 display(null.mlm)
 summary(null.mlm)
 
 # test our model against the null model
 anova(null.mlm, mlm4)


```
